A saturation model for impaired learning with enhanced plasticity: Supplementary material

### 2013-09-21 MHCpaper/MHCsupp.tex
pooled model for dep only

% -*- TeX -*- -*- UK -*-
% ----------------------------------------------------------------
\documentclass[12pt]{article}
%Preamble:
\usepackage{a4wide}
\input{sl_preamble.tex}
\hypersetup{colorlinks,linkcolor=blue,citecolor=teal,urlcolor=teal}
\input{sl_graphics_preamble.tex}
\graphicspath{{"Figures/"}}
%
\usepackage{array}
\newcounter{tablerow}
\newcommand{\tablerowautorefname}{row}
\newenvironment{tabularn}[1]{\setcounter{tablerow}{0}\begin{tabular}{#1>{\refstepcounter{tablerow}}l}}{\end{tabular}}
%\usepackage{etoolbox}
%\AtBeginEnvironment{tabular}{\setcounter{tablerow}{0}}
%
% ----------------------------------------------------------------
% New commands etc.
\input{sl_definitions.tex}
\input{sl_symbols.tex}
%matrices
\newcommand{\I}{\mathbf{I}}
%vec of ones
%equilibrium distribution
\newcommand{\pr}{\mathbf{p}}
\newcommand{\eq}{\pr^\infty}
%other symbols
\newcommand{\w}{\mathbf{w}}
\newcommand{\W}{\mathbf{W}}
\newcommand{\frg}{\W^\mathrm{F}}
\newcommand{\M}{\mathbf{M}}
\newcommand{\F}{\boldsymbol{\Phi}}
\newcommand{\pot}{^{\text{pot}}}
\newcommand{\dep}{^{\text{dep}}}
\newcommand{\potdep}{^{\text{pot/dep}}}
\newcommand{\norm}{_0}
\newcommand{\inc}{_{\text{inc}}}
\newcommand{\dec}{_{\text{dec}}}
\newcommand{\incdec}{_{\text{inc/dec}}}
\newcommand{\wt}{_{\text{WT}}}
\newcommand{\ko}{_{\text{D$^\mathrm{b}$K$^\mathrm{b}$-/-}}}
\newcommand{\KO}{D$^\mathrm{b}$K$^\mathrm{b}$-/-}
\newcommand{\tpre}{t_{\text{pre}}}
\newcommand{\ttrain}{t_{\text{train}}}
\newcommand{\lmax}{_{\text{max}}}
\newcommand{\lmin}{_{\text{min}}}
% ----------------------------------------------------------------
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Title info:
\title{A saturation model for impaired learning with enhanced plasticity: Supplementary material}
%
%% Author List:
%%
%\author{An Author$^a$, Another Author$^b$ and Yet Another Author$^c$\\
%%
%\small{\emph{$^a$ Address 1}}\\
%\small{\emph{$^b$ Address 2}}\\
%\small{\emph{$^c$ Address 3}}\\
%}

@@

This model is parameterised by the range of values, $q\in[q\lmin,q\lmax]$, for potentiation and depression.
We will use the same values for potentiation and depression in the wild-type as well as potentiation in the \KO\ mutant.
We will use larger values for depression in the \KO\ mutant.
The values of these parameters are listed in \autoref{tab:params}.

@@

\begin{table}
 \begin{center}
  \begin{tabularn}{|l|c|c|c|c|c|c|}
    \cline{1-7}
    % after \\: \hline or \cline{col1-col2} \cline{col3-col4} ...
    Model & \# states & pot,WT dep & \KO\ dep & $\Delta f$ & $r\ttrain$ & $r\tpre$ \\
    \cline{1-7}
    Serial        & 10 & $q=0.3$  & $q=0.4$  & 0.1  & 5 & 20  &\label{tr:multistate_weak} \\
    Serial        & 10 & $q=0.3$  & $q=0.4$  & 0.3  & 5 & 20  &\label{tr:multistate_med} \\
    Serial        & 10 & $q=0.3$  & $q=0.4$  & 0.45 & 5 & 30  &\label{tr:multistate_strong} \\
    Two-state     & 2  & $q=0.1$  & $q=0.2$  & 0.1  & 5 & 5   &\label{tr:binary} \\
    Multistate    & 10 & $q=0.3$  & $q=0.4$  & 0.3  & 5 & 5   &\label{tr:multistate_lin} \\
    Pooled        & 10 & $q\in[0.3,0.4]$  & $q\in[0.6,0.8]$
                                          & 0.1 & 5 & 20 &\label{tr:pooled_plenty}\\
    Pooled        & 10 & $q\in[0.05,0.4]$  & $q\in[0.1,0.8]$
                                          & 0.1 & 5 & 20 &\label{tr:pooled_scarce}\\
    Cascade       & 10 & $x=0.25$ & $x=0.33$ & 0.3  & 5 & 20  &\label{tr:cascade_short} \\
    Cascade       & 10 & $x=0.25$ & $x=0.33$ & 0.3  & 5 & 100 &\label{tr:cascade_long} \\
    Non-uniform MS       & 10 & $x=0.25$ & $x=0.33$ & -0.3  & 50 & 150 &\label{tr:nonuni} \\
    \cline{1-7}
%   \hline
%    % after \\: \hline or \cline{col1-col2} \cline{col3-col4} ...
%    & Model & \# states & pot,WT dep & \KO\ dep & $\Delta f$ & $r\ttrain$ & $r\tpre$ \\
%    \cline{1-7}
%    \label{tr:cascade_short}&
%    Cascade    & 10 & $x=0.25$ & $x=0.33$ & -0.3 & 50 & 50   \\
%    \label{tr:cascade_long}&
%    Cascade    & 10 & $x=0.25$ & $x=0.33$ & -0.3 & 50 & 150  \\
%    \label{tr:multistate_weak}&
%    Multistate & 10 & $q=0.6$  & $q=0.8$  & -0.1 & 50 & 50   \\
%    \label{tr:multistate_strong}&
%    Multistate & 10 & $q=0.6$  & $q=0.8$  & -0.4 & 20 & 20   \\
%    \label{tr:binary}&
%    Two-state  & 2  & $q=0.4$  & $q=0.8$  & -0.1 & 5  & 5    \\
%    \label{tr:pooled_plenty}&
%    Pooled     & 10 & $q\in[0.3,0.4]$  & $q\in[0.6,0.8]$
%                                          & -0.1 & 70 & 70 \\
%    \label{tr:pooled_scarce}&
%    Pooled     & 10 & $q\in[0.1,0.4]$  & $q\in[0.2,0.8]$
%                                          & -0.1 & 70 & 70 \\
%    \cline{1-7}
  \end{tabularn}
 \end{center}
  \caption{Parameters used in simulations.} \label{tab:params}
\end{table}

@@

@@

There are two effects here as well.
Smaller $\beta$ will increase the probability of crossing the centre of the chain, speeding up learning, but it will also concentrate probability at the ends of the chain, depleting the centre and slowing down learning (see \autoref{fig:multistate_strong}\ref{fig:multistate_strong_eq}\ref{fig:multistate_strong_eq_KO}, blue trace).
The first effect goes like $1/\beta$, whereas the second goes like $\beta^{M/2}$ and will be more significant for smaller $\beta$ or in a longer chain.

@@

@@

@@

@@

@@

@@

>>>>

% -*- TeX -*- -*- UK -*-
% ----------------------------------------------------------------
\documentclass[12pt]{article}
%Preamble:
\usepackage{a4wide}
\input{sl_preamble.tex}
\hypersetup{colorlinks,linkcolor=blue,citecolor=teal,urlcolor=teal}
\input{sl_graphics_preamble.tex}
\graphicspath{{"Figures/"}}
%
\usepackage{array}
\newcounter{tablerow}
\newcommand{\tablerowautorefname}{row}
\newenvironment{tabularn}[1]{\setcounter{tablerow}{0}\begin{tabular}{#1>{\refstepcounter{tablerow}}l}}{\end{tabular}}
%\usepackage{etoolbox}
%\AtBeginEnvironment{tabular}{\setcounter{tablerow}{0}}
%
% ----------------------------------------------------------------
% New commands etc.
\input{sl_definitions.tex}
\input{sl_symbols.tex}
%matrices
\newcommand{\I}{\mathbf{I}}
%vec of ones
%equilibrium distribution
\newcommand{\pr}{\mathbf{p}}
\newcommand{\eq}{\pr^\infty}
%other symbols
\newcommand{\w}{\mathbf{w}}
\newcommand{\W}{\mathbf{W}}
\newcommand{\frg}{\W^\mathrm{F}}
\newcommand{\M}{\mathbf{M}}
\newcommand{\F}{\boldsymbol{\Phi}}
\newcommand{\pot}{^{\text{pot}}}
\newcommand{\dep}{^{\text{dep}}}
\newcommand{\potdep}{^{\text{pot/dep}}}
\newcommand{\norm}{_0}
\newcommand{\inc}{_{\text{inc}}}
\newcommand{\dec}{_{\text{dec}}}
\newcommand{\incdec}{_{\text{inc/dec}}}
\newcommand{\wt}{_{\text{WT}}}
%\newcommand{\ko}{_{\text{D$^\mathrm{b}$K$^\mathrm{b}$-/-}}}
%\newcommand{\KO}{D$^\mathrm{b}$K$^\mathrm{b}$-/-}
\newcommand{\ko}{_{\text{MHC}}}
\newcommand{\KO}{MHC}
\newcommand{\tpre}{t_{\text{pre}}}
\newcommand{\ttrain}{t_{\text{train}}}
\newcommand{\lmax}{_{\text{max}}}
\newcommand{\lmin}{_{\text{min}}}
% ----------------------------------------------------------------
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Title info:
\title{A saturation model for impaired learning with enhanced plasticity: Supplementary material}
%
%% Author List:
%%
%\author{An Author$^a$, Another Author$^b$ and Yet Another Author$^c$\\
%%
%\small{\emph{$^a$ Address 1}}\\
%\small{\emph{$^b$ Address 2}}\\
%\small{\emph{$^c$ Address 3}}\\
%}

@@

This model is parameterised by the range of values, $q\in[q\lmin,q\lmax]$, for potentiation and depression.
We will use the same values for potentiation and depression in the wild-type as well as potentiation in the \KO\ mutant.
We will use larger values for depression in the \KO\ mutant.
In the main text, we consider a version of this model that has resource depletion for depression only, so that potentiation transition probabilities are unaffected by the number of potentiated synapses.
This is done by setting $q\pot\lmax=q\pot\lmin$.
The values of these parameters are listed in \autoref{tab:params}.

@@

\begin{table}
 \begin{center}
  \begin{tabularn}{|l|c|c|c|c|c|c|}
    \cline{1-7}
    % after \\: \hline or \cline{col1-col2} \cline{col3-col4} ...
    Model & \# states & pot & WT dep & \KO\ dep & $\Delta f$ & $r\tpre$ \\
    \cline{1-7}
    Serial        & 10 & $q=0.3$  & $q=0.3$  & $q=0.4$  & 0.1  & 20  &\label{tr:multistate_weak} \\
    Serial        & 10 & $q=0.3$  & $q=0.3$  & $q=0.4$  & 0.3  & 20  &\label{tr:multistate_med} \\
    Serial        & 10 & $q=0.3$  & $q=0.3$  & $q=0.4$  & 0.45 & 30  &\label{tr:multistate_strong} \\
    Two-state     & 2  & $q=0.1$  & $q=0.1$  & $q=0.2$  & 0.1  & 5   &\label{tr:binary} \\
    Multistate    & 10 & $q=0.3$  & $q=0.3$  & $q=0.4$  & 0.3  & 5   &\label{tr:multistate_lin} \\
    Pooled        & 10 & $q\in[0.3,0.4]$  & $q\in[0.3,0.4]$  & $q\in[0.6,0.8]$
                                          & 0.1 & 20 &\label{tr:pooled_plenty}\\
    Pooled        & 10 & $q\in[0.05,0.4]$ & $q\in[0.05,0.4]$  & $q\in[0.1,0.8]$
                                          & 0.1 & 20 &\label{tr:pooled_scarce}\\
    Pooled        & 7  & $q=0.008$        & $q\in[0.0006,0.6]$  & $q\in[0.001,1]$
                                          & 0.4 & 20 &\label{tr:pooled_deponly}\\
    Cascade       & 10 & $x=0.25$ & $x=0.25$ & $x=0.33$ & 0.3  & 20  &\label{tr:cascade_short} \\
    Cascade       & 10 & $x=0.25$ & $x=0.25$ & $x=0.33$ & 0.3  & 100 &\label{tr:cascade_long} \\
    Non-uniform MS& 10 & $x=0.25$ & $x=0.25$ & $x=0.33$ & 0.3  & 150 &\label{tr:nonuni} \\
    \cline{1-7}
%   \hline
%    % after \\: \hline or \cline{col1-col2} \cline{col3-col4} ...
%    & Model & \# states & pot,WT dep & \KO\ dep & $\Delta f$ & $r\ttrain$ & $r\tpre$ \\
%    \cline{1-7}
%    \label{tr:cascade_short}&
%    Cascade    & 10 & $x=0.25$ & $x=0.33$ & -0.3 & 50 & 50   \\
%    \label{tr:cascade_long}&
%    Cascade    & 10 & $x=0.25$ & $x=0.33$ & -0.3 & 50 & 150  \\
%    \label{tr:multistate_weak}&
%    Multistate & 10 & $q=0.6$  & $q=0.8$  & -0.1 & 50 & 50   \\
%    \label{tr:multistate_strong}&
%    Multistate & 10 & $q=0.6$  & $q=0.8$  & -0.4 & 20 & 20   \\
%    \label{tr:binary}&
%    Two-state  & 2  & $q=0.4$  & $q=0.8$  & -0.1 & 5  & 5    \\
%    \label{tr:pooled_plenty}&
%    Pooled     & 10 & $q\in[0.3,0.4]$  & $q\in[0.6,0.8]$
%                                          & -0.1 & 70 & 70 \\
%    \label{tr:pooled_scarce}&
%    Pooled     & 10 & $q\in[0.1,0.4]$  & $q\in[0.2,0.8]$
%                                          & -0.1 & 70 & 70 \\
%    \cline{1-7}
  \end{tabularn}
 \end{center}
  \caption{Parameters used in simulations.} \label{tab:params}
\end{table}

@@

This impaired learning after pre-training in the wild-type is caused by the excessive potentiation pushing the synapses away from the central transition that generates the learning signal.
Essentially, this model has a form of metaplasticity where repeated potentiation makes subsequent depression harder.
This allows excessive pre-training to impair learning, despite increasing the number of potentiated synapses.

@@

There are two effects here as well.
Smaller $\beta$ will increase the probability of crossing the centre of the chain, speeding up learning, but it will also concentrate probability at the ends of the chain, depleting the centre and slowing down learning (see \autoref{fig:multistate_strong}\ref{fig:multistate_strong_eq}\ref{fig:multistate_strong_eq_KO}, blue trace).
The first effect goes like $1/\beta$, whereas the second goes like $\beta^{M/2}$ and will be more significant for smaller $\beta$ or in a longer chain.
This exponential decay in the initial distribution is what allows the saturation bias to overwhelm the increases intrinsic plasticity rate and lead to impaired learning with enhanced plasticity.
It is important that a learning signal is only generated by the central transition so that the exponential decay can take effect.

@@

Intuitively, this model is missing two features of the \hyperref[sec:multistate]{multistate model}.
First, it does not have the exponential amplification of the effect of initial saturation bias as this model does not have a chain of states for the distribution to decay across.
Second, it does not have the metaplastic effect where repeated potentiation makes future depression harder, as potentiation will merely increase the number of potentiated synapses.

@@

Like the \hyperref[sec:binary]{two state model}, this model is missing two features of the \hyperref[sec:multistate]{multistate model}.
First, it does not have the amplification of the effect of initial saturation bias as every transition contributes to the learning signal, so the exponentially decaying distribution has no effect.
Second, it does not have the metaplastic effect where repeated potentiation makes future depression harder, as potentiation will merely increase the number of potentiated synapses without pushing them away from any boundary between strong and weak states.

@@

\resultsfig{pooled resource}{unaffected potentiation}{($q\pot\lmax=q\pot\lmin$)}{pooled_deponly}

@@

The situation discussed in the main text, with resource depletion for depression only and potentiating transitions  unaffected, is shown in \autoref{fig:pooled_deponly}.
It shows the same learning patterns as the more symmetric versions.

This model shares one feature of the \hyperref[sec:multistate]{multistate model}.
The effect of initial saturation bias is amplified by resource depletion, allowing for impaired learning with enhanced plasticity.
However, it does not have the metaplastic effect where repeated potentiation makes future depression harder, in fact it shows the opposite effect due to resource replenishment, so it can never show impaired learning after pre-training.

@@

This model shares both features of the \hyperref[sec:multistate]{multistate model}.
First, the effect of initial saturation bias is amplified by the exponential decay of transition probabilities away from the middle.
Second, it has the metaplastic effect where repeated potentiation makes future depression harder, also due to he exponential decay of transition probabilities.

@@

Like the \hyperref[sec:cascade]{cascade model}, this model also shares both features of the \hyperref[sec:multistate]{multistate model}.
First, the effect of initial saturation bias is amplified by the exponential decay of transition probabilities away from the middle.
Second, it has the metaplastic effect where repeated potentiation makes future depression harder, also due to he exponential decay of transition probabilities.
